---
title: "Hw03"
output: github_document
---

```{r}      
library(dplyr)
library(ggplot2)
library(scales)
dat <- read.csv('~/stat133/stat133-hws-fall17/hw03/data/nba2017-teams.csv')
```

# Ranking of Teams
 


```{r}
#NBA Teams Ranked by Total Salary
ggplot(dat, aes(reorder(team, salary), salary)) +
         geom_bar(stat = "identity") +
         labs(title = "NBA Teams Ranked by Total Salary", x = "Team", y = "Salary") +
         coord_flip() +
         geom_hline(aes(yintercept =  mean(salary)), col = "red")
```

```{r}
#Nba Teams ranked by Total Points
ggplot(dat, aes(reorder(team, points), points)) +
         geom_bar(stat = "identity") +
         labs(title = "NBA Teams ranked by Total Points", x = "Team", y = "Points") +
         coord_flip() +
         geom_hline(aes(yintercept = 14 * mean(points)), col = "red")
```

```{r}
#rankings by Efficiency
ggplot(dat, aes(reorder(team, efficiency), efficiency)) +
         geom_bar(stat = "identity") +
         labs(title = "NBA Teams Ranked by Efficiency", x = "Team", y = "EFF") +
         coord_flip() +
         geom_hline(aes(yintercept = 14 * mean(efficiency)), col = "red")
```

The first plot doesn't give much insight into the team rankings because salary doesn't have a direct effect in winning or losing games. However, teams that spend more are more likely to have better talent, and would therefore do better.

The second plot is a little more informative, as points scored directly affects a teams chance of winning games. 

The third plot seems like it is the most informative, as it takes into account many different statistics covering all facets of the game.

### PCA Rankings

```{r}
pca_stats <- dat %>% 
  select(points3_made, points2_made, points1_made, off_rebounds, 
         def_rebounds, assists, steals, blocks, turnovers, fouls)

pca <- prcomp(pca_stats, scale. = TRUE)
eigenvalue = pca$sdev^2
prop = round(eigenvalue/sum(eigenvalue), 4)
 cumprop = round(cumsum(prop), 4)
eig <- data.frame(eigenvalue = eigenvalue, 
                  prop = prop,
                  cumprop = cumprop
                  )
eig

pc1 <- pca$x[ ,1]
pc2 <- pca$x[ ,2]
dat$pc1 <- pc1
dat$pc2 <- pc2
```

```{r}
#Scatterplot
ggplot(dat, aes(x = pc1, y = pc2)) +
         geom_point() +
         labs(title = "PCA plot (PC1 and PC2)", x = "PC1", y = "PC2") +
         geom_text(aes(label = dat$team)) +
         geom_hline(yintercept = 0) +
         geom_vline(xintercept = 0)
```

```{r}
#Index Based on PC1

s1 <- 100 * ((pc1 - min(pc1))/(max(pc1) - min(pc1)))
dat$s1 <- s1

#barplot of PC1
ggplot(dat, aes(reorder(team, s1), s1)) +
  labs(title = "NBA Teams ranked by scaled PC1", x = "Team", y = "First PC (scaled from 0 to 100") +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_hline(yintercept = 14 * mean(dat$s1))
  


```

The PC1 index provides a measure of a team's ability based on various statistics

## Comment & Reflection

It was and I think it was very tough and I really struggled with it. I used relative paths before to try it out on the midterm and they are important because they allow me to access files from current directory. I have used R script before but it was good practice. The r script part was difficult and nothing was too easy. It took about 7 hours to do this and preparing the data took the longest.






